---
title: "Understanding GDP Relations to Crime in US Cities"
author: "Kai Yoshino, Isabella Spaletta, Chrisopther Ross, Marisa Nanakul"
date: "March 6, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(rmarkdown)
source("resources.R")
```

### What Are The Datasets?

The two datasets that we used in this anaylsis are one for crime based data on U.S. cities over a 5 year time period (2010-2015) and one for the GDP for metropolitan areas in the U.S. over a 5 year time period (2010-2015). 

#### FBI Crime Dataset

Below is an example of what the crime dataset looks like (shows all 5 years): 

```{r, echo=FALSE}
  kable(cities.all[1:5,])
```

##### Observations Made About the Dataset
* The size of the dataset is {rows, columns}:

```{r, echo=FALSE}
  dim(cities.all)
```

* Column Names and Data Types of the crime table:


```{r, echo=FALSE}
  sapply(cities.all, typeof)
```


##### Other Observations & Changes Made to Dataset
* All crime datasets were individually captured by year, but we combined them to make one large crime dataset.
* If a city has a crime value of NA, this means this data could not be found, not that any of these crimes did not occur. NA does not equal 0 in this dataset.
* Violent crimes is defined of four offenses: murder and nonnegligent manslaughter, rape, robbery, and aggravated assault.
* Property crimes is defined of four offenses: burglary, larceny theft, motor vehicle theft, and arson.
<br>
<br>

#### GDP Dataset

```{r, echo=FALSE}
  gdp[1:5,] %>% 
  select_("X2010_gdp", "X2011_gdp", "X2012_gdp", "X2013_gdp", "X2014_gdp", "X2015_gdp", "gdp_per_person_2010", "gdp_per_person_2011",	"gdp_per_person_2012",	"gdp_per_person_2013", "gdp_per_person_2014",	"gdp_per_person_2015") %>% 
  kable()
```

##### Observations Made About the Dataset
* The size of the dataset is {rows, columns}:

```{r, echo=FALSE}
  dim(gdp %>% 
  select_("X2010_gdp", "X2011_gdp", "X2012_gdp", "X2013_gdp", "X2014_gdp", "X2015_gdp", "gdp_per_person_2010", "gdp_per_person_2011",	"gdp_per_person_2012",	"gdp_per_person_2013", "gdp_per_person_2014",	"gdp_per_person_2015"))
```

* Column Names and Data Types of the crime table:

```{r, echo=FALSE}
  colnames(gdp %>% 
  select_("X2010_gdp", "X2011_gdp", "X2012_gdp", "X2013_gdp", "X2014_gdp", "X2015_gdp", "gdp_per_person_2010", "gdp_per_person_2011",	"gdp_per_person_2012",	"gdp_per_person_2013", "gdp_per_person_2014",	"gdp_per_person_2015"))
```

```{r, echo=FALSE}
  sapply(gdp %>% 
  select_("X2010_gdp", "X2011_gdp", "X2012_gdp", "X2013_gdp", "X2014_gdp", "X2015_gdp", "gdp_per_person_2010", "gdp_per_person_2011",	"gdp_per_person_2012",	"gdp_per_person_2013", "gdp_per_person_2014",	"gdp_per_person_2015"), typeof)
```


### What We Found
The main purpose of combining these two data sets was to determine if there was some sort of correlation between a cities Gross Domestic Product, their population and the amount of crime that happens within it.  Now when initially looking at a data set like this it is very easy to just used the hard numbers to determine if there is any correlation between these three things.  If for example you were to look at just the population of a city vs the number of crimes committed in that city you would find that, why yes they are related. And if you were to plot this data it would look something like this: 

_Note: The population data has been passed through a *log* function which basically compresses the data so that data that was closer is easily visible and those data points that were outliers are now in a closer grouping._

``` {r echo=FALSE, message=FALSE, warning=FALSE}
total.crime.vs.pop <- ggplot(gdp, aes(x=log(Population_2015), y=total_crime_2015, color="red")) +
                    geom_point() + 
                    geom_smooth(col = "black") +
                    ggtitle("Population and Total Crimes in 2015") +
                    labs(x = "Log(Population)", y= "Total Crimes Committed")

total.crime.vs.pop
```

While it makes sense that as Populuation increases the number of crimes increase, but we believed that just that information is not what should be the focus when looking at trends in crime.  Instead we created a value that we named *Crime Rate* that is the total crimes committed in a city divided by the population.  This new value is the total number of crimes per person within a city, or what we refer to as the crime rate.  _Note:_ This is not to be confused with the more common term "crime rate" that refers to the number of crimes committed per 100,000 people. 
<br> 
<br> 

###Crime Rate with GDP and Population

So what we did is we took the newly made variable *Crime Rate* and compared it with the population of the different cities and their Gross Domestic Product per Person. If we break all that down here are how the different data points relate to each other
``` {r echo=FALSE, message=FALSE, warning=FALSE}
colnames(time.poisson.data)[4] <- "Year"

time.poisson.data$Year <- gsub("Population_", "", time.poisson.data$Year)

crime.vs.pop.2015 <- ggplot(time.poisson.data, aes(x=population_value, y=person_crime_value, color=Year)) +
                  geom_point(shape=1) +
                  stat_smooth(method = "glm", col = "black") +
                  ggtitle("Population and Crimes commited per person over 2010-2015") +
                  labs(x = "Population", y= "Crime Rates")

gdp.vs.crime <- ggplot(time.poisson.data, aes(x=person_gdp_value, y=person_crime_value, color=Year)) +
                  geom_point(shape=1) +
                  stat_smooth(method = "glm", col = "black") +
                  ggtitle("Crimes per Person and GDP per person over 2010-2015") +
                  labs(x = "GDP per Person", y= "Crime Rate")

gdp.vs.pop.2015 <- ggplot(time.poisson.data, aes(x=population_value, y=person_gdp_value, color=Year)) +
                  geom_point(shape=1) +
                  stat_smooth(method = "glm", col = "black") +
                  ggtitle("Population and GDP per person over 2010-2015") +
                  labs(x ="Population", y="GDP Per Person")
              
gdp.vs.pop.2015

```

```{r echo=FALSE, warning=FALSE, message=FALSE}
crime.vs.pop.2015
```

```{r echo=FALSE, warning=FALSE, message=FALSE}
gdp.vs.crime
```

Each of the graphs has been given a regression line to make it easier to see the different trends that occur within each of the comparissons. Yet to truly understand if there is a statistical signifigance to each of these we needed to run a multiple variable Poisson Regression.  A Poisson Regression allows us to work with count data and data that is not distributed normally.  Meaning when you look at the graph it doesn't look like a bell.  As the graphs demonstrate this is very much true for our data set.  After running our Poisson Regression we came up with this equation: 

#### Crime Rate = 4.938e-7(GDP Per Person) - 5.72e-8(Population) - 3.086

What this breaks down to is that for every one dollar that GDP per person increase the Crime rate increase by 0.0000004938 and for every person added to the population crme rate decreases by 0.0000000527.  Theses numbers seem incredibly small but when you are talking about millions of dollars in GDP and thousands of people it becomes increasingly important.  After computing this equation it was still found to not be statistically signifigant, meaning that the GDP per Person and the Population had no real impact on the Crime Rate. 

### Crime Rate Change 

After looking at just the crime rate over the 5 year time period we created a new variable which was the difference in crime rate, GDP per person, and Population change between the years.  At this point we had started to work more with countinous data than count data, so we decided to use a standard linear regression to find the equation and signifigance of the data.  

``` {r echo=FALSE, message=FALSE, warnings=FALSE}
rate.crime.vs.pop <- ggplot(tmp, aes(x=population_change_value, y=total_crime_rate_change_value)) +
                  geom_point(shape=1) +
                  stat_smooth(method = "lm", col = "red") +
                  ggtitle("Population and Total Crime Rate Change") +
                  labs(x ="Population", y="Total Crime Rate Change")
rate.crime.vs.pop
```

``` {r echo=FALSE, message=FALSE, warnings=FALSE}
rate.crime.vs.pop <- ggplot(tmp, aes(x=gdp_rate_change_value, y=total_crime_rate_change_value)) +
                  geom_point(shape=1) +
                  stat_smooth(method = "lm", col = "red") +
                  ggtitle("GDP Rate Change and Total Crime Rate Change") +
                  labs(x ="GDP Rate Change", y="Total Crime Rate Change")
rate.crime.vs.pop
```

This is where the data got very interesting.  As we can see there is a very signifigant change on the Difference in the rate of crime based on the population in a negative fashion.  This means that every time the population difference increase the rate of crime decreases.  But on the other side of thins the GDP change actually increases the Crime change.  Both of these factors were statistically signifigant, meaning the both affect the difference in the crime rate.  

### What this all means. 


